Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-occurrence Where the Hudsons's Bay Lowlands meets the Ontario Shield

### 2024-04-09 reports/GnC_report/paper/paper.qmd
ptogress

These two features are often better addressed by tools not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners. Indeed, machine learning datasets can be very diverse and often more detailed than camera trap datasets, requiring processing tools with more detailed features. Tagging is a common step in machine learning research and in industry, and if often referred to as labeling.

@@

In order to reduce the impact of false positives on our workflow, we resorted to adding a pre-processing step in our workflow in which all images were manually sorted into "wildlife" and "nonwildlife" folders inside each subfolders. In addition, because wildtrax does not track the upload path of an image, and instead tracks the unique name of each file, we are required to change the way we encode unique images. Therefore, we developed a simple and very efficient script in the Julia programming language to paralellize the renaming of each file into a standardized manner. We had originally refrained from using this method of uniquely identifying files and instead relied on tracking the unique path of each file as we wanted to reduce the impact on our storage capacity (as it essentially required to double the required storage space).

@@

Wildtrax keeps track of image batches using "location folders". We created our own location folders but combining images within the same retrieval, site and location IDs. For example the image mentioned above would be placed in the "TC1_P028_P028-1_WLU-10" location folders. For locations in which images were previously sorted into "wildlife" and "nonwildlife", we declined each location folders into a "_W" and "_NW" variants and uploaded as such to wildtrax.

@@

sum_tab <- tibble::tribble(
  ~Status,                ~Images,           ~Detections,
  "Total",                nb_image_fmt,      dets_tb_cat_unfil_sum_fmt,
  "With detections",      dets_tb_fmt[1],    "N/A",
  "Empty",                dets_tb_fmt[2],    "N/A",
  "Filtered (> 0.1)",     uploaded_fmt,      dets_tb_cat_sum_fmt,
  "Processed",            processed_img_fmt, processed_fmt
)

@@

\newpage

# Appendix I: Species table

@@

# Appendix II: Land cover classes table

@@

# Appendix III: List of camera trapping tools compared

<!-- The following line inserts a page break  -->
\newpage

>>>>

These two features are often better addressed by tools not traditionally used for tagging camera trap images, but used by the larger community of machine learning practitioners. Indeed, machine learning datasets can be very diverse and often more detailed than camera trap datasets, requiring processing tools with more detailed features (for example, see @Schneider2021). Tagging is a common step in machine learning research and in industry, and if often referred to as labeling.

@@

Wildtrax keeps track of image batches using "location folders". We created our own location folders by combining images within the same retrieval, site and location IDs. The standard format we used for location folders follows this convention:

- "\textcolor{red}{retrievalID}\_\textcolor{blue}{siteID}\_\textcolor{green}{locationID}\_\textcolor{purple}{cameraID}

Giving, for example, a folder with the name:

- "\textcolor{red}{TC1}\_\textcolor{blue}{P028}\_\textcolor{green}{P028-1}\_\textcolor{purple}{WLU-10}"

For locations in which images were previously sorted into "wildlife" and "nonwildlife", we declined each location folders into a "_W" and "_NW" variants and uploaded as such to wildtrax.

In order to reduce the impact of false positives on our workflow, we resorted to adding a pre-processing step in our workflow in which most of the images were manually sorted into "wildlife" and "nonwildlife" folders inside each of the camera subfolders. In addition, because wildtrax does not track the upload path of an image, and instead tracks the unique name of each file, we are required to change the way we encode unique images. Therefore, we developed a simple and very efficient script in the Julia programming language to paralellize the renaming of each file into a standardized manner. We had originally refrained from using this method of uniquely identifying files and instead relied on tracking the unique path of each file as we wanted to reduce the impact on our storage capacity (as it essentially required to double the required storage space).

@@

@@

sum_tab <- tibble::tribble(
  ~Status,                ~Images,           ~Detections,               ~`False Positives`,
  "Total",                nb_image_fmt,      dets_tb_cat_unfil_sum_fmt, "N/A",
  "With detections",      dets_tb_fmt[1],    "N/A",                     "N/A",
  "Empty",                dets_tb_fmt[2],    "N/A",                     "N/A",
  "Filtered (> 0.1)",     uploaded_fmt,      dets_tb_cat_sum_fmt,       "N/A",
  "Processed",            processed_img_fmt, processed_fmt,             dels_nb_fmt
)

@@

# Appendices

## Appendix I: Species table

@@

## Appendix II: Land cover classes table

@@

## Appendix III: List of camera trapping tools compared

-  Aardwolf
-  Agouti
-  AIDE
-  animl
-  animl (R)
-  CAIMAN
-  Camelot
-  Camera Base
-  Camera Trap Manager
-  CameraTrapDetectorR
-  CameraTraps
-  Camtrap DP
-  camtrapR
-  Conservation AI
-  Cos4Cloud FASTCAT
-  CPW Photo Warehouse
-  DeepFaune
-  digiKam
-  EcoAssist
-  eMammal
-  Event Finder Suite
-  IBEIS
-  IL2BB BBoxEE
-  Mapview
-  mbaza
-  MD Fast API
-  MLWIC MLWIC2 
-  PhotoSpread
-  rapidPop
-  Renamer & CamTrap
-  Snoopy
-  sparcD
-  SpeedyMouse
-  Timelapse
-  TrailCamData
-  TRAPPER
-  U-Infuse
-  ViXeN
-  Wild ID
-  WILDCAM
-  wildcam
-  Wildepod
-  WildEye TrapTagger
-  Wildlife Insights
-  Wildlife observer network
-  Wildme Wildbook
-  WildTrax
-  WTB Where's The Bear
-  zamba
-  Zooniverse
